From Reversible Programs to \\ Univalent Universes and Back

### 2017-05-12 upi.tex
fixed pi2
figure fibration
better fib figure and some direction for sec 3

\[\begin{array}{rcl}
\tau &::=& \bt \\
\\
v &::=& \begin{array}[t]{lrcl}
                    & \fc &:& \bt \\
              \alt & \tc &:& \bt
               \end{array} \\
\\
c &::=& \begin{array}[t]{lrcl}
              & \id &:& \tau \iso \tau \\
               \alt & \swap &:& \bt \iso \bt \\
               \alt & \circ &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_3)
                              \to (\tau_1 \iso \tau_3)
               \end{array} \\
\\
! &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_1) \\
\invc{\id} &=& \id \\
\invc{\swap} &=& \swap \\
\invc{(\compc{c_1}{c_2})} &=& \compc{\invc{c_2}\;}{\;\invc{c_1}} \\
\\
\alpha &::=& \begin{array}[t]{lrcl}
               & \id &:& c \isotwo c \\
               \alt & \assocc &:& \compc{c_1}{(\compc{c_2}{c_3})} \isotwo
                                              \compc{(\compc{c_1}{c_2})}{c_3} \\
               \alt & \idlc &:& \compc{\id}{c} \isotwo c \\
               \alt & \idrc &:& \compc{c}{\id} \isotwo c \\
               \alt & \invl &:& \compc{c\;}{\;\invc{c}} \isotwo \id \\
               \alt & \invr &:& \compc{\invc{c}}{c} \isotwo \id \\
               \alt & \bullet &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_3)
                                            \to (c_1 \isotwo c_3) \\
               \alt & \odot &:& (c_1 \isotwo c_1') \to (c_2 \isotwo c_2')
                                            \to (\compc{c_1}{c_2} \isotwo \compc{c_1'}{c_2'}) \\
             \end{array}
\end{array}\]

The 2-combinators $\alpha$ also have inverses. The above forms a 2-groupoid.

>>>>

\[\begin{array}{rcl}
\tau &::=& \bt \\
\\
v &::=& \begin{array}[t]{lrcl}
                    & \fc &:& \bt \\
              \alt & \tc &:& \bt
               \end{array} \\
\\
c &::=& \begin{array}[t]{llcl}
              & \id &:& \tau \iso \tau \\
               \alt & \swap &:& \bt \iso \bt \\
               \alt & ! &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_1) \\
               \alt & \circ &:& (\tau_1 \iso \tau_2) \to (\tau_2 \iso \tau_3) \to (\tau_1 \iso \tau_3)
               \end{array} \\
\\
\alpha &::=& \begin{array}[t]{lrcl}
               & \id &:& c \isotwo c \\
            \alt & \idlc &:& \compc{\id}{c} \isotwo c \\
            \alt & \idrc &:& \compc{c}{\id} \isotwo c \\
            \alt & \invl &:& \compc{c\;}{\;\invc{c}} \isotwo \id \\
            \alt & \invr &:& \compc{\invc{c}}{c} \isotwo \id \\
            \alt & \rho &:& \swap \circ \swap \isotwo \id \\
            \alt & \assocc &:& \compc{(\compc{c_1}{c_2})}{c_3} \isotwo \compc{c_1}{(\compc{c_2}{c_3})} \\
            \alt & \odot &:& (c_1 \isotwo c_1') \to (c_2 \isotwo c_2') \to (\compc{c_1}{c_2} \isotwo \compc{c_1'}{c_2'}) \\
            \alt & !! &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_1) \\
            \alt & \bullet &:& (c_1 \isotwo c_2) \to (c_2 \isotwo c_3) \to (c_1 \isotwo c_3)
             \end{array}
\end{array}\]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Univalent Type Families}

We will consider a general statement of \emph{univalence} for type families. We
start with a some preliminaries. Fig.~\ref{fig:fib} depicts a general type
family $P : A \rightarrow \mathcal{U}$ mapping each point $x : A$ to a space
$P(x)$. As everything is continuous, we expect this type family to respect
identities. In other words if $x:A$ and $y:A$ are identified, it must be the
case that the spaces $P(x)$ and $P(y)$ are ``equivalent.''

\begin{figure}
\begin{tabular}{c@{\qquad\qquad}c}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $A$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$x$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$y$};
  \draw (-1,-2) ellipse (0.5cm and 2cm);
  \node[left] at (-1.5,-2) {Fiber $P(x)$};
  \draw (1,-2) ellipse (0.5cm and 2cm);
  \node[right] at (1.5,-2) {Fiber $P(y)$};
\end{tikzpicture}
&
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (0,-5) ellipse (2cm and 0.8cm);
  \node[below] at (0,-6) {Base Space $A$};
  \draw[fill] (-1,-5) circle [radius=0.025];
  \node[below] at (-1,-5) {$x$};
  \draw[fill] (1,-5) circle [radius=0.025];
  \node[below] at (1,-5) {$y$};
  \draw (-1.3,-2) ellipse (0.5cm and 2cm);
  \node[left] at (-1.8,-2) {Fiber $P(x)$};
  \draw (1.3,-2) ellipse (0.5cm and 2cm);
  \node[right] at (1.8,-2) {Fiber $P(y)$};
  \draw[below,cyan,thick] (-1,-5) -- (1,-5);
  \node[below,cyan,thick] at (0,-5) {$p$};
  \draw[->,red,ultra thick] (-0.8,-1.7) to [out=45, in=135] (0.8,-1.7);
  \draw[->,red,ultra thick] (0.8,-2.3) to [out=-135, in=-45] (-0.8,-2.3);
  \node[red,ultra thick] at (0,-2) {$\simeq$};
\end{tikzpicture}
\end{tabular}
\caption{\label{fig:fib}(left)Type family $P : A \rightarrow \mathcal{U}$ as a
  fibration with total space $\Sigma_{(x:A)} P(x)$; (right) a path $x \equiv y$
  induces an equivalence between the spaces $P(x)$ and $P(y)$}
\end{figure}

The converse statement is whether an equivalence between $P(x)$ and $P(y)$
necessarily induces an identification $x \equiv y$. It is easy to construct
several examples in which this converse statement fails and it instructive to
examine such examples.
